Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2015-04-07 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Added IWBG_NG graphs and added equations still need some formatting aid

According to the United States Geological Survey, most gypsum in the United States is used to make wallboards for homes and other structures.  A typical home in the United States contains more than 7 metric tons of gypsum ~\cite{USGS:2015}.  As defined by ASTM, gypsum wallboard is gypsum board designed for use on walls, ceilings, or partitions and that affords a surface suitable to receive decoration~\cite{ASTM_C1396}. The ASTM standard further describes gypsum wallboard as having a noncombustible core, surfaced with paper bonded to the core. By weight, wallboard is composed of more than 85~\% gypsum, less than 10~\% cellulose, less than 3~\% starch and approximately 1~\% of both crystalline silica and fibrous glass~\cite{GA:2010}. Chemically gypsum can be described as calcium sulfate dihydrate, (CaSO$_{4}$2H$_{2}$O) .  By weight pure gypsum is approximately 79~\% calcium sulfate and 21~\% chemically bound water~\cite{Lawson:1977}.

@@

Figures~\ref{IWGB_NG_Temp_FrontSet},\ref{IWGB_NG_Temp_RearCenter} and \ref{IWGB_NG_HF_RearCenter} show the values for the natural gas burner experiments with the gypsum wallboard.  Revise to match Ch 3 set

@@

 \begin{figure}[p]
 	\centering
 	\includegraphics[width=3.5in]{../Figures/IWGB_NG_TC_Surface_Center_Avg}\\
 	\includegraphics[width=3.5in]{../Figures/IWGB_NG_TC_Surface_Offset_Avg}\\
 	\includegraphics[width=3.5in]{../Figures/IWGB_NG_TC_Plume_Avg}\\
 	\caption[Averaged peak wall and plume temperatures for the natural gas burner]{Averaged peak wall and plume temperatures for the natural gas burner fires at several different positions from the instrumented gypsum board wall.  The positions ranged from 0.3 m away from the wall to physically touching the wall, 0.0 m.}
 	\label{IWGB_NG_Temp_FrontSet}
 \end{figure}

@@

@@

\subsection{Correlation Descriptions}

@@

>>>>

According to the United States Geological Survey, most gypsum in the United States is used to make wallboards for homes and other structures.  A typical home in the United States contains more than 7 metric tons of gypsum ~\cite{USGS:2015}.  As defined by ASTM, gypsum wallboard is gypsum board designed for use on walls, ceilings, or partitions and that affords a surface suitable to receive decoration~\cite{ASTM_C1396}. The ASTM standard further describes gypsum wallboard as having a noncombustible core, surfaced with paper bonded to the core. By weight, wallboard is composed of more than 85~\% gypsum, less than 10~\% cellulose, less than 3~\% starch and approximately 1~\% of both crystalline silica and fibrous glass~\cite{GA:2010}. Chemically gypsum can be described as calcium sulfate dihydrate, (CaSO$_{4}$ $\bullet$ 2H$_{2}$O) .  By weight pure gypsum is approximately 79~\% calcium sulfate and 21~\% chemically bound water~\cite{Lawson:1977}.

@@

Figures~\ref{IWGB_NG_TC_Plume_set},\ref{IWGB_NG_TC_Wall_set} and \ref{IWGB_NG_TC_Backside_set} show the values for the natural gas burner experiments with the gypsum wallboard.

@@

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGBNG01_TC_Plume_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGBNG04_TC_Plume_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGBNG05_TC_Plume_Avg}\\
	\caption[Plume temperatures for the natural gas fueled burner]{Plume temperatures along the centerline of the natural gas fueled burner.    Each graph displays data from a different distance between the edge of the burner and the instrumented gypsum wallboard.  Beginning with the top graph and moving down, the distances from the wall are 0.31~m, 0.15~m, and 0~m.}
	\label{IWGB_NG_TC_Plume_set}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGBNG01_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGBNG04_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGBNG05_TC_Surface_Center_Avg}\\
	\caption[Front surface wall temperatures for the natural gas fueled burner]{Front surface wall temperatures for the natural gas fueled burner. Each graph displays data from a different distance between the edge of the burner and the instrumented gypsum wallboard.  Beginning with the top graph and moving down, the distances from the wall are 0.31~m, 0.15~m, and 0~m.}
	\label{IWGB_NG_TC_Wall_set}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGBNG01_TC_Back_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGBNG04_TC_Back_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGBNG05_TC_Back_Center_Avg}\\
	\caption[Backside wall temperatures for the natural gas fueled burner]{Backside wall temperatures along the centerline of the natural gas fueled burner. Each graph displays data from a different distance between the edge of the burner and the instrumented gypsum wallboard.  Beginning with the top graph and moving down, the distances from the wall are 0.31~m, 0.15~m, and 0~m.}
	\label{IWGB_NG_TC_Backside_set}
\end{figure}

 \begin{figure}[p]
 	\centering
 	\includegraphics[width=3.5in]{../Figures/IWGBNG01_HF_Center_Avg}\\
 	\includegraphics[width=3.5in]{../Figures/IWGBNG04_HF_Center_Avg}\\
 	\includegraphics[width=3.5in]{../Figures/IWGBNG05_HF_Center_Avg}\\
 \caption[Wall heat flux along the centerline of the natural gas fueled burner]{Wall heat flux along the vertical centerline of the natural gas fueled burner. Each graph displays data from a different distance between the edge of the burner and the instrumented gypsum wallboard.  Beginning with the top graph and moving down, the distances from the wall are 0.31~m, 0.15~m, and 0~m.}
 	\label{IWGB_NG_HF_CenterSet}
 \end{figure}

 \begin{figure}[p]
 	\centering
 	\includegraphics[width=3.5in]{../Figures/IWGBNG01_HF_Offset_Avg}\\
 	\includegraphics[width=3.5in]{../Figures/IWGBNG04_HF_Offset_Avg}\\
 	\includegraphics[width=3.5in]{../Figures/IWGBNG05_HF_Offset_Avg}\\
 	\caption[Wall heat flux along the edge of the natural gas fueled burner]{Wall heat flux along the edge of the natural gas fueled burner. Each graph displays data from a different distance between the edge of the burner and the instrumented gypsum wallboard.  Beginning with the top graph and moving down, the distances from the wall are 0.31~m, 0.15~m, and 0~m.}
 	\label{IWGB_NG_HF_EdgeSet}
 \end{figure}

@@

Ratio from heat flux from source fire data 
vyto second item ignite

$\dot{q}'' = (\dot{Q} \chi_r)/(4 \pi\ R^2) $ \\

$\dot{q}''$ = heat flux per unit area $(kW/m^2)$ \\
$\dot{Q}$ = heat release rate $(kW)$ \\
$ \chi_r$ = radiative fraction \\
$R$ = distance from the point source to the target $(m)$ \\

Source fire plume temp and shape factor from vertical cylinder

$\dot{q}'' = \epsilon \sigma F_12 T^4 $ \\

$\epsilon$ = emissivity  \\
$\sigma$ = Stefan-Boltzmann Constant $(5.670373(21)×10^(−8) W/m^2 K^4)$  \\

@@

discuss the premise  
McCaffery

$Z_c = 0.08 dot{Q}^(2/5)$ \\

$Z_i = 0.20 dot{Q}^(2/5)$ \\

$Z_c$ = continuous flame height $(m)$  \\

$Z_i$ = intermittent flame height $(m)$ \\

Heskestad widely used  NRC CFITrainer  texts

\subsection{Correlation Descriptions}
Heskestad, developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion.  Appropriate fuels would “include liquid pool fires and other horizontal surface fires” add cites.  Given that the fuel configurations are square, a conversion of the fuel surface area to the effective diameter is required in order to use Heskestad’s flame height equation directly.  Setting the area of the fuel surface, 0.093~$m^2$, equal to π $D^2$/4 and then solving for $D$ yields an effective diameter of 0.344~m.
Calculating the flame heights with the average peak heat release rates measured from the three fuels yields 1.0 m (3.3 ft), 1.0 m (3.3 ft), and 0.73 m (2.4 ft) for natural gas, gasoline, and polyurethane foam respectively. Taking the uncertainty of the measurements used to determine the heat release rate (±11~\%) into account generates a range of values.   The range of calculated values are compared with the  mean flame height and range for flame heights  based on photographic analysis for each of the fuels in Table z.

\begin{center} 
$ H_f = 0.235\: \dot{Q}^{(2/5)} - 1.02 D $  \\
\end{center}
$ H_f$ = flame height (m)  \\
$ \dot{Q}$ = heat release rate (kW) \\
$ D$ = diameter of burner (m)  \\

@@

Compare Natural Gas at 0.61 m from instrumented wall and at 0.0 m from instrumented wall - problem bounds.
